Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-07-11 doc/manuscript/manuscript.Rmd
overhaul results section

The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
The forward-backward selection process revealed severity, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the total variation.
ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.033 (P = 0.001).
Assessment of the marginal effects showed that state, host, and year significantly explained genetic variation (P = 0.001), but aggressiveness was only mildly significant (P = 0.045). 
Much of the variation appears to be driven by isolates from Mexico (Fig. \@ref(rda-biplot)) and 2005; it's worth noting that all Mexican isolates were collected in 2005.
Variance partitioning of the independent variables indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.

@@

We observed 87 mycelial compatibility groups (MCG), the most abundant of which represented 73 isolates over 37 multilocus genotypes (MLG) (Fig. \@ref(barplot). 
Our assessment of aggressiveness association with MCGs showed a significant effect (P < 0.001) indicating that MCG 44 is, on average, more aggressive than other MCGs (Table \@ref(tab:mcg-aggressiveness)).

@@

We observed a total of 165 multilocus genotypes (MLG) (215 with 16 loci).
The most abundant MLG represented 27 isolates (Table \@ref(tab:mlg-table)) across five regions (NE, MI, WA, CO, and ND). 
The genotypes were relatively evenly distributed within regions with moderate to high diversity (Table \@ref(tab:iatab)). 
Of the 165 MLGs, 76 (46%) were found across all geographic regions except WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
Our graph-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
Community analysis showed 51 communities, 23 of which consisted of a single MLG/MCG pair. 
The three communities with the most members contained 8 of the ten most abundant MCGs. 
Overall, however, there was no clear pattern to the association between MLG and MCGs.

@@

These six cultivars had a total of 238 *S. sclerotiorum* isolates (of 366 total isolates collected) belonging to 130 MLGs.
Of these MLGs, 53 were singletons, while the rest were obtained from more than one host cultivar; only one private allele on locus 55-4 was identified, belonging to the *S. sclerotiorum* population from the cultivar 'Bunsi'.
The five most abundant MLGs with associated MCGs are shown in Table \@ref(tab:mlg-table).
The majority of MLGs reported were found on multiple host cultivars and only a few unique to a cultivar.
For instance, there were only 15 unique MLGs within the population of 68 *S. sclerotiorum* isolates obtained from the 'Beryl' cultivar.

@@

Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig. \@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig. \@ref(community-graph16)).
The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87 x 10^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.

## DAPC {-}

![
Discriminant Analysis on regions showing that Mexico is differentiated from other populations. 
**(a)** Mean population assignment (horizontal axis) probability (vertical axis) for all populations with N > 10 (facets).
Numbers next to source populations indicate population size.
All values sum to one.
**(b)** Scatter plot of first two components from DAPC.
Points represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
The center of each component is represented as grid lines.
](../../results/figures/publication/DAPC.pdf){#DAPC width=100%}

DAPC analysis for different regions and hosts are depicted in Fig. 8.
These isolates were clone-corrected at location/host/region/year level prior to the analysis (N =318).
Accordingly, MX (N = 15) and CA (N =18) populations grouped separately from all other populations with MX isolates showing the highest deviation (Fig. 8A).
Comparison of genotypes by region showed all of the nine MLGs in the Mexican population were unique, with no genotypes crossing over to other populations.
MLGs of Californian isolates were found only in OR and WA populations.
MLG 6, 84 and 100 were (see Supplementary Table 1 for SSR genotypes related to 165 MLGs of total isolates) unique to CA isolates.

Populations from different regions were compared with AMOVA, where differentiation was low among regions ($\Phi_{PT}$  = 0.11; P = 0.001).
Pairwise comparisons of all populations according to region showed MX and CA populations were significantly different across all populations at P $\leq$ 0.001.
The $\Phi_{PT}$ values of Pairwise comparisons between CA and other populations ranged from 0.116 to 0.389 and from 0.355 to 0.493 between MX and other population.
The CA population showed evidence of sexual <!-- ZNK: There is the sample size caveat here that should be discussed. -->
recombination ($\bar{r}_d$ = 0.008; P = 0.27, $I_A$ = 0.08; P = 0.3) while the $\bar{r}_d$ value of AU was 0.12 with a high P of 0.05 ($I_A$ = 0.70; P = 0.05), whereas all other populations showed linkage disequilibrium for loci tested indicating high possibility of clonal reproduction.

The DAPC for hosts was limited to six most common cultivars: G122 (N = 55), Beryl (N = 62), Bunsi (N = 47), Yellow (N = 26), Pinto (N = 15), and Merlot (N = 11).
According to Fig. 8B, cultivar Merlot isolates group distant from others, but were not significantly different (P =0.325).  Interestingly, *S. sclerotiorum* populations from cultivars 'Beryl' and 'G122', which are known to have different level of white mold susceptibility, showed no difference in population genetic data.

AMOVA revealed genetic differences among isolates sampled in different years (DAPC biplot not shown).
The $\Phi_{PT}$ was a low value of 0.07 with P = 0.001.
Isolates from 2012 were not included in the analysis since there were only 4 samples.
Samples for other years ranged from 19 in 2009 to 63 in 2005.
Sampling year 2003 to 2005 and 2009 were different across every Pairwise comparison.
None of the above sampling variables (i.e. region, year, host) seemed to influence distinct molecular composition of *S. sclerotiorum* populations since pre-defined (i.e. region, host, and year) isolates and isolates that were assigned into post groups by DAPC varied largely (Figs. 8A, 8B and 9).
However, assignment of isolates by DAPC analysis of regions did show some promising results for MX and CA isolates (Fig. 8A).
The original group size of each region and DAPC defined group sizes for different regions are given in Table 6.

-------------------------------------------------------------------------------
Hierarchy               d.f.        S.S.       % variation    $\Phi statistic$  
-------------------- ---------- ------------- -------------- ------------------
Between Source         1 (1)    0.897 (0.753) -0.158 (0.116)   0.136 (0.0965)

@@

Since `find.clusters()` function of *adegenet* did not result in a clear-cut number of clusters for 318 isolates, we arrived at six (6) as the best number of populations (Fig. 9) for our data after examining results of DAPC for different number of clusters.
Three clusters, namely clusters one, three, and six were formed separately from the rest.
The cluster 3 of the DAPC plot consisted entirely of MX isolates (N =13/15).
Rest of the two MX isolates were grouped in cluster 1.
Cluster 6 had 50 isolates in total representing mainly of NE (N = 15/37), MI (N = 12/58), and CO (N = 12/34).
All CA (N = 18) isolates and 13/58 WA isolates dominated the cluster 1.
AMOVA on these six clusters resulted 44% molecular variance among populations with P = 0.001.
Pairwise comparison showed significant difference (P = 0.001) for all pairs.
The DAPC plot after pooling all isolates together and removing clones (N =165) was compared with the plot of 318 isolates looked similar to Fig. 8C.
The number of clusters was set to 6 for the analysis which resulted in a DAPC having similar topology to plot in Fig. 9 with 318 isolates(results not shown).
Isolates of MX grouped separately while another distinct group had 24 isolates representing CO, NE, and ND.

We tested the clone corrected data for evidence of sexual outcrossing within *S. sclerotiorum* populations in different regions and different years/region.
CA showed evidence of sexual outcrossing (Ia = 0.08, P.Ia = 0.317, $\bar{r}_d$ = 0.008, P.$\bar{r}_d$ = 0.317) while all other regions implied clonal propagation (data not shown).
Five populations showed evidence of sexual outcrossing: CA in 2004, FR in 2004 and 2005, OR in 2003, and WA in 2008.
However, after clone-correction these populations had 6-7 isolates/population **(Table 7)** and therefore, may not be sufficient to make inferences about reproductive mode.
Only exception was the WA_2008 population, which consisted of 12 isolates.
Within sampling locations white mold nurseries (wmn) were tested for linkage disequilibrium since same wmn fields were used for sampling over different years.
Only CA wmn showed evidence of sexual reproduction ( = 0.008; P = 0.29) while all other wmn locations suggested clonal reproduction.
The DAPC analyses of WA, and MI isolates from same wmn fields over different years are in Fig. 10.
Since *S. sclerotiorum* isolates are predominantly clonal, we wanted to explore whether there was any introduction of new isolates from other fields or if there were evidence of sexual reproduction.
The AMOVA results for different clusters of Fig. 10A revealed significance difference between WA 2008 isolates from the rest ($\Phi_{PT}$ $\geq$ 0.206, P = 0.001).
The 2003 and 2005 isolates were genetically similar ($\Phi_{PT}$ = 0.056, P = 0.129).
MI isolates from 2008 and 2003 (Fig. 10B) showed significant difference ($\Phi_{PT}$ = 0.127, P = 0.04) while 2005 isolates did not differ molecularly from both 2008 and 2003 isolates ($\Phi_{PT}$ $\leq$ 0.078, P $\geq$ 0.076).
NE wmn isolates sampled in 2004 and 2008 were genetically similar ($\Phi_{PT}$ = 0.042, P = 0.179) (DAPC results not shown).
The standardized index of association of wmn isolates from each region (i.e.
MI, WA, and NE) for above mentioned sampling years suggested clonal reproduction ( $\geq$ 0.05, P $\leq$ 0.002).

>>>>

The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
The forward-backward selection process revealed severity, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the total variation.
ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.033 (P = 0.001).
Assessment of the marginal effects showed that state, host, and year significantly explained genetic variation (P = 0.001), but aggressiveness was only mildly significant (P = 0.045). 
Much of the variation appears to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)); it's worth noting that all Mexican isolates were collected in 2005.
Variance partitioning of the independent variables indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.

@@

We observed 87 mycelial compatibility groups (MCG), the most abundant of which represented 73 isolates over 37 multilocus genotypes (MLG) (Fig. \@ref(barplot)). 
Our assessment of aggressiveness association with MCGs showed a significant effect (P < 0.001) indicating that MCG 44 is, on average, more aggressive than other MCGs (Table \@ref(tab:mcg-aggressiveness)).

@@

@@

We observed a total of 165 multilocus genotypes (MLG) (215 with 16 loci).
The most abundant MLG represented 27 isolates (Table \@ref(tab:mlg-table)) across five regions (NE, MI, WA, CO, and ND). 
The genotypes were relatively evenly distributed within regions with moderate to high diversity (Table \@ref(tab:iatab)). 
Of the 165 MLGs, 76 (46%) were found across all geographic regions except WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig. \@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig. \@ref(community-graph16)).
The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87 x 10^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.

Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
Community analysis showed 51 communities, 23 of which consisted of a single MLG/MCG pair. 
The three communities with the most members contained 8 of the ten most abundant MCGs. 
Overall, however, there was no clear pattern to the association between MLG and MCGs.

@@

## Population Differentiation {-}

### Analysis of Molecular Variance

The analysis of molecular variance (AMOVA) over the hierarchy of Source and Region showed no significant variation between white mold nurseries and non-white mold nurseries after accounting for variation between regions (Table \@ref(tab:amova)).
Because Mexico shared no genotypes with any of the other populations and we suspected that it was contributing to some of the observed variance, we re-ran AMOVA without the Mexican isolates and found similar decreased, but still significant levels of differentiation between regions.
In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no evidence of population differentiation (Supplementary Data).

-------------------------------------------------------------------------------
Hierarchy               d.f.        S.S.       % variation    $\Phi$ statistic  
-------------------- ---------- ------------- -------------- ------------------
Between Source         1 (1)    0.897 (0.753) -0.158 (0.116)   0.136 (0.0965)

@@

### Discriminant Analysis of Principal Components

Discriminant Analysis of Principal Components (DAPC) was performed with the first 21 principal components, representing 88% of the total variance. 
The first discriminant axis (representing 64% of the discriminatory power) separated the centroid for the Mexico isolates from the rest of the data, indicating strong differentiation (Fig. \@ref(DAPC)b). 
The second discriminant axis separated the centroid for the Californian isolates.
The mean population assignment probabilities for all populations with n > 10 showed that only isolates from Mexico, CA, and France had > 50% probabilities of being reassigned to their source populations (Fig. \@ref(DAPC)a).

![
Discriminant Analysis of Principal Components on regions showing that Mexico is differentiated from other populations. 
**(a)** Mean population assignment (horizontal axis) probability (vertical axis) for all populations with N > 10 (facets).
Numbers next to source populations indicate population size.
All values sum to one.
**(b)** Scatter plot of first two components from DAPC.
Points represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
The center of each component is represented as grid lines.
](../../results/figures/publication/DAPC.pdf){#DAPC width=100%}

### Linkage Analysis {-}

Nearly all populations showed evidence for linkage (\@ref(fig:iatab)), which serves as evidence for clonal reproduction or other forms of non-random mating. 
The only exceptions are CA and Australia. Both of these populations showed only moderate significance (P < 0.05) with $\bar{r}_d$ values of 0.03 and 0.12, respectively.
